The five pillars of computational reproducibility: Bioinformatics and beyond

### 2023-06-12 manuscript/5pillars.Rmd
removed the docker image because it belongs to docker inc

Containers are an attempt to resolve some of the downsides of VMs.
Container images can be thought of as similar to VMs, but are more lightweight as they
share parts of the host operating system [@Nust2020-cr].
In the example shown in **Figure 4**, the six containerised applications share the same operating
system, while the three VM applications each involve their own operating system which incurs a
significant performance overhead @noauthor_2021-lt.
Therefore, running workflows in containers incurs only a small reduction in performance as compared
to running directly on the host system @Di_Tommaso2015-pw.
Container images are highly portable because containers include everything needed to run the
application, including the system tools and libraries, ensuring that the environment is consistent
across different systems.
For example, the most popular containerisation system, Docker [@Merkel2014-ek], makes it
possible to run Windows and Linux/Unix containers on any computer with Docker installed, with the
promise of reproducibility.
While there are several alternatives such as Podman, Rkt or Apptainer/Singularity, Docker remains the most widely used
containerisation system for web development and research.
Due to its early adoption, Docker now has a large community of users, extensive documentation and
a vast collection of pre-built container images in the DockerHub registry.
Docker can fetch images from DockerHub and run them on the host system with just a couple of commands,
and typically within a few minutes.
This accelerates the installation procedure dramatically which is a known bottleneck for bioinformatic
reproducibility [@Mangul2019-ei;@Mangul2019-xp].

![Figure 4. Architecture of containers (left) compared to VMs (right) @noauthor_2021-lt.](../img/docker-containerized-and-vm-transparent-bg.webp)

Another potential solution to this problem is to use a package/environment management system such as
[Conda](https://conda.org/) or [Guix](https://guix.gnu.org/).
These allow users to create, manage, and deploy software packages,
dependencies, and environments across different computing platforms @Perkel2023-cv.
Conda was initially developed to solve the problem of package dependency management in Python,
but it now supports many other programming languages, including R, C/C++, Java, and others.
Conda allows researchers to create isolated environments with specific versions of
packages, so users can have different versions of R or Python in different environments
on the same host system.
Conda environments have limitations around portability and consistency, as such environments
are tied to specific operating systems and hardware architecture.

>>>>

Containers are an attempt to resolve some of the downsides of VMs.
Container images can be thought of as similar to VMs, but are more lightweight as they
share parts of the host operating system [@Nust2020-cr].
In the example shown in **Figure 4**, the five containerised applications share the same operating
system, while the three VM applications each involve their own operating system which incurs a
significant performance overhead @noauthor_2021-lt.
Therefore, running workflows in containers incurs only a small reduction in performance as compared
to running directly on the host system @Di_Tommaso2015-pw.
Container images are highly portable because containers include everything needed to run the
application, including the system tools and libraries, ensuring that the environment is consistent
across different systems.
For example, the most popular containerisation system, Docker [@Merkel2014-ek], makes it
possible to run Windows and Linux/Unix containers on any computer with Docker installed, with the
promise of reproducibility.
While there are several alternatives such as Podman, Rkt or Apptainer/Singularity, Docker remains the most widely used
containerisation system for web development and research.
Due to its early adoption, Docker now has a large community of users, extensive documentation and
a vast collection of pre-built container images in the DockerHub registry.
Docker can fetch images from DockerHub and run them on the host system with just a couple of commands,
and typically within a few minutes.
This accelerates the installation procedure dramatically which is a known bottleneck for bioinformatic
reproducibility [@Mangul2019-ei;@Mangul2019-xp].

```{r,containers,fig.cap="Figure 4. Architecture of containers (left) compared to VMs (right). Adapted from @noauthor_2021-lt.",echo=FALSE,fig.height=6}

library("RColorBrewer")
cols <- brewer.pal(5, "Set2")
cols2 <- adjustcolor( cols, alpha.f = 0.7)
plot(0,type='n',axes=FALSE,ann=FALSE, xlim=c(0, 12), ylim=c(0, 8))

#border
polygon(c(0,12,12,0),c(0,0,8,8),border="gray")

polygon(c(0,5.5,5.5,0),c(0,0,0.9,0.9), col=cols2[1], border = NA)
polygon(c(0,5.5,5.5,0),c(1,1,1.9,1.9), col=cols2[2], border = NA)
polygon(c(0,5.5,5.5,0),c(2,2,2.9,2.9), col=cols2[3], border = NA)

polygon(c(0,1,1,0),c(3,3,5,5), col=cols2[4], border = NA)
polygon(c(1.1,2.1,2.1,1.1),c(3,3,5,5), col=cols2[4], border = NA)
polygon(c(2.2,3.2,3.2,2.2),c(3,3,5,5), col=cols2[4], border = NA)
polygon(c(3.3,4.3,4.3,3.3),c(3,3,5,5), col=cols2[4], border = NA)
polygon(c(4.4,5.4,5.4,4.4),c(3,3,5,5), col=cols2[4], border = NA)

labs1 <- c("Hardware","Host Operating System","Container Engine")
text(c(2.7,2.7,2.7),c(0.4,1.4,2.4),labels=labs1)

labs2 <- c("App 1","App 2","App 3","App 4","App 5")
text(c(0.5,1.6,2.7,3.8,4.9) ,c(4,4,4,4,4), labels=labs2,srt=90)

lines(c(0,0,5.5,5.5),c(5.2,5.5,5.5,5.2)  )
text(2.7,5.7,labels="Containerized applications")

polygon(c(6,11.5,11.5,6),c(0,0,0.9,0.9), col=cols2[1], border = NA)
polygon(c(6,11.5,11.5,6),c(1,1,1.9,1.9), col=cols2[3], border = NA)

labs3 <- c("Hardware","Hypervisor")
text(c(8.5,8.5),c(0.4,1.4),labels=labs3)

polygon(c(6,7.8,7.8,6),c(2,2,4.5,4.5), col=cols2[2], border = NA)
polygon(c(8,9.6,9.6,8),c(2,2,4.5,4.5), col=cols2[2], border = NA)
polygon(c(9.8,11.5,11.5,9.8),c(2,2,4.5,4.5), col=cols2[2], border = NA)

text(c(6.85,8.8,10.65),c(3.5,3.5,3.5),labels="Guest\nOperating\nSystem")

polygon(c(6,7.8,7.8,6),c(4.5,4.5,6,6), col=cols2[4], border = NA)
polygon(c(8,9.6,9.6,8),c(4.5,4.5,6,6), col=cols2[4], border = NA)
polygon(c(9.8,11.5,11.5,9.8),c(4.5,4.5,6,6), col=cols2[4], border = NA)

text(c(6.85,8.8,10.65),c(5,5,5),labels=labs2[1:3])

lines(c(6,6,11.5,11.5),c(6.2,6.5,6.5,6.2)  )
text(8.5,6.7,labels="Virtualized applications")

```
Another potential solution to this problem is to use a package/environment management system such as
[Conda](https://conda.org/) or [Guix](https://guix.gnu.org/).
These allow users to create, manage, and deploy software packages,
dependencies, and environments across different computing platforms @Perkel2023-cv.
Conda was initially developed to solve the problem of package dependency management in Python,
but it now supports many other programming languages, including R, C/C++, Java, and others.
Conda allows researchers to create isolated environments with specific versions of
packages, so users can have different versions of R or Python in different environments
on the same host system.
Conda environments have limitations around portability and consistency, as such environments
are tied to specific operating systems and hardware architecture.
